---
title: "Introduction on how to use Select package to find Evolutionary Dependencies."
author: "Arvind Iyer"
email: "arvind.iyer@unil.ch"
date: "`r Sys.Date()`"
output: 
 rmarkdown::html_vignette:
    css: doc.css
vignette: >
  %\VignetteIndexEntry{introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

- The goal of select pacakge is to detect Evolutionary Dependecies between alterations/genes in cancer.
- `select` package provides function to generate the backgorund model and other utilites functions.

### Installation

- You can install the development version of select from [GitHub](https://github.com/CSOgroup/select) with:
``` r
# install.packages("devtools")
# devtools::install_github("CSOgroup/select")
```

### Example

- We will run select algorithm on processed LUAD dataset from TCGA provided with the package.


```{r}
library(select)
library(dplyr)
```

#### Data Description & Format

- The loaded data is list object which consists of 
  - gam: a presence absence matrix of alterations
  - samples: a named vector of sample annotations
  - alt: a named vector of alteration annotations

```{r}
# Check the data strucutre
## Load the data provided with the package
data(luad_data, package = "select")
str(luad_data)
```

#### Running select

- We use the function `select()` which generates the background model and results.
- The parameters for the functions are:
  - `M`: the list object of GAMs & TMB
  - `sample.class`: a named vector of samples with covariates
  - `alteration.class`: a named vector of alteration with covariates
  - many more

- The function returns a dataframe.

```{r}
alpi <- select::select(
  M = luad_data$gam,
  sample.class = luad_data$samples,
  alteration.class = luad_data$alt,
  folder = './',
  r.seed = 110,
  n.cores = 1,
  vetos = NULL,
  n.permut = 100,
  min.feature.support=5,
  min.feature.freq=0.001,
  remove.0.samples=TRUE,
  remove.unknown.class.samples=TRUE,
  rho = 0.1,
  lambda = 15,
  save.intermediate.files = FALSE,
  randomization.switch.threshold = 30,
  max.memory.size=100,
  calculate_APC_threshold = TRUE,
  calculate_FDR = TRUE,
  verbose = TRUE
  )
```

#### Intrepreting the results
- Lets look into the results
- The below table explains what each coloum means.

| Colnames                                       | Meaning                                  |
|------------------------------------------------|------------------------------------------|
| SFE_1                                          | Selected Functional Event (SFE_1)        |
| SFE_2                                          | Selected Functional Event  (SFE_2)       |
| name                                           | Interaction Motif                        |
| type_1                                         | Type of mutation of SFE_1                |
| type_2                                         | Type of mutation of SFE_2                |
| int_type                                       | Interaction Motif  type                  |
| support_1                                      | Samples mutated with  SFE_1              |
| support_2                                      | Samples mutated with  SFE_2              |
| freq_1                                         | Frequency of SFE_1                       |
| freq_2                                         | Frequency of SFE_2                       |
| overlap                                        | Co-mutation between   SFE_1 and SFE_2    |
| max_overlap                                    | Maximum possible   Co-mutation           |
| freq_overlap                                   | Frequency of   Co-mutation               |
| r_overlap                                      | Background   Co-mutation                 |
| r_freq_overlap                                 | Background frequency   of Co-mutation    |
| diff_overlap                                   | Difference of   co-mutations             |
| abs_diff_overlap                               | Absoulte Difference   of co-mutations    |
| direction                                      | Interaction Type                         |
| wMI_stat                                       | Weighted Mutual   Information            |
| wMI_p.value                                    | P-value on Mutual   Information          |
| ME_p.value                                     | P-value on   co-mutation                 |
| E.r.wMI_stat                                   | Background Weighted   Mutual Information |
| MI_diff                                        | Difference of Mutual   Information       |
| wMI_p.value_FDR                                | FDR                                      |
| select_score_good_cancer_cell_2017_criterion_1 | Cancer Cell Paper Criteria               |
| select_score                                   | Effect Size (select Score)               |


```{r}
options(width=100)
# Look into dataframe
(alpi %>% filter(wMI_p.value_FDR) %>% arrange(desc(select_score))) %>% head(2) 
# Total significant Hits
alpi  %>% filter(wMI_p.value_FDR) %>% count(wMI_p.value_FDR,direction)
```


## Suggestion

- Please go through the methods on best way to filter and use the method as mentioned in the paper.
  - Mina, M., Iyer, A., Tavernari, D., Raynaud, F., & Ciriello, G. (2020). Discovering functional evolutionary dependencies in human cancers. Nature Genetics, 52(11), 1198-1207. 


### SessionInfo
```{r}
# Print the sessionInfo
sessionInfo()
```